Examples in R* paper

### 2020-03-19 src/s_examples.Rmd
added Aki's suggestions about exploring wider data

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(caret)
library(gbm)
library(mvtnorm)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
source("monitornew.R")
```

@@

Use gbm model in Caret to predict chain identity
```{r}
library(caret)
library(gbm)

# generate data
full_data <- f_generate_lower_var_four(1/3, 0.3, 1, 1000) %>% 
  melt() %>% 
  rename(iter=Var1,
         chain=Var2) %>% 
  dplyr::select(-iter)

@@

caretGrid <- expand.grid(interaction.depth=c(3), n.trees = 50,
                   shrinkage=c(0.1),
                   n.minobsinnode=10)
gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid)
# plot(gbmFit1)
plda <- predict(object=gbmFit1, newdata=testing_data)
tibble(predicted=plda, actual=testing_data$chain) %>% 
  mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
  summarise(mean(correct))

@@

>>>>

```{r}
rm(list=ls())
library(tidyverse)
library(reshape2)
library(rstan)
library(latex2exp)
library(caret)
library(gbm)
library(mvtnorm)
options(mc.cores=4)
rstan_options(auto_write = TRUE)
source("monitornew.R")

caretGrid <- expand.grid(interaction.depth=c(3), n.trees = 50,
                   shrinkage=c(0.1),
                   n.minobsinnode=10)
```

@@

Use gbm model in Caret to predict chain identity
```{r}
# generate data
full_data <- f_generate_lower_var_four(1/3, 0.3, 1, 1000) %>% 
  melt() %>% 
  rename(iter=Var1,
         chain=Var2) %>% 
  dplyr::select(-iter)

@@

gbmFit1 <- train(chain ~ ., data = training_data, 
                 method = "gbm",
                 trControl = trainControl(method = 'none'), 
                    tuneGrid = caretGrid)
# plot(gbmFit1)
plda <- predict(object=gbmFit1, newdata=testing_data)
tibble(predicted=plda, actual=testing_data$chain) %>% 
  mutate(correct=if_else(predicted==actual, 1, 0)) %>% 
  summarise(mean(correct))

@@

## Aki Vehtari's suggestion: try measure for similar numbers of iterations to posterior
Centered
```{r}
N <- 250
A <- readRDS("../output/A_matrix.rds")
model <- stan_model("mvt_250.stan")
fit <- sampling(model, data=list(N=N, A=A), iter=500, chains=4, thin=1)
saveRDS(fit, "../output/mvt_fit_500.rds")
fit <- readRDS("../output/mvt_fit_500.rds")

full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)
m_flattened <- matrix(nrow = nrow(full_data) * 4, ncol = 252)
k <- 1
for(i in 1:4){
  for(j in 1:nrow(full_data)){
    m_flattened[k, 1:251] <- full_data[j, i, ]
    m_flattened[k, 252] <- i
    k <- k + 1
  }
}
m_flattened <- m_flattened %>% 
  as.data.frame() %>% 
  rename(chain=V252) %>% 
  mutate(chain=as.factor(chain))

a_df <- tibble(samples_cp=colMeans(mAccuracy)/(1/n_distinct(training_data$chain)))

mon <- monitor(fit)
mean(mon$Rhat>1.01)
mean(mon$Bulk_ESS<400)
mean(mon$Tail_ESS<400)
```

Non-centered
```{r}
A <- readRDS("../output/A_matrix.rds")
model <- stan_model("mvt_250_ncp.stan")
fit <- sampling(model, data=list(N=N, A=A), iter=500, chains=4, thin=1)
saveRDS(fit, "../output/mvt_ncp_fit_500.rds")

full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)
nparams <- dim(full_data)[3]
m_flattened <- matrix(nrow = nrow(full_data) * 4, ncol = (nparams + 1))
k <- 1
for(i in 1:4){
  for(j in 1:nrow(full_data)){
    m_flattened[k, 1:nparams] <- full_data[j, i, ]
    m_flattened[k, (nparams + 1)] <- i
    k <- k + 1
  }
}
m_flattened <- m_flattened %>% 
  as.data.frame() %>% 
  rename(chain=V502) %>% 
  mutate(chain=as.factor(chain))
n_distinct(m_flattened$chain)

a_df <- a_df %>%
  mutate(samples_ncp=colMeans(mAccuracy)/(1/n_distinct(training_data$chain)))
saveRDS(a_df, "../output/mvt_500_all.rds")
g <- 
  a_df %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8, bins=20) +
  scale_fill_grey("Model", labels=c("centered", "noncentered")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  theme(text = element_text(size=14, colour="black"),
        legend.position = c(0.8, 0.8))
ggsave("../output/wide_data.pdf", g, width = 12, height = 8)
```

# Wide dataset - standard normal
```{r}
N <- 10000
A <- diag(nrow=N)
model <- stan_model("mvt_wide.stan")
fit <- sampling(model, data=list(N=N), iter=400, chains=4)
saveRDS(fit, "../output/mvt_wide_400.rds")

full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)
nparams <- dim(full_data)[3]
m_flattened <- matrix(nrow = nrow(full_data) * 4, ncol = (nparams + 1))
k <- 1
for(i in 1:4){
  for(j in 1:nrow(full_data)){
    m_flattened[k, 1:nparams] <- full_data[j, i, ]
    m_flattened[k, (nparams + 1)] <- i
    k <- k + 1
  }
}
m_flattened <- m_flattened %>% 
  as.data.frame()
colnames(m_flattened)[nparams + 1] <- "chain"
m_flattened <- m_flattened %>% 
  mutate(chain=as.factor(chain))
n_distinct(m_flattened$chain)

a_df <- tibble(samples_400=colMeans(mAccuracy)/(1/n_distinct(training_data$chain)))
saveRDS(a_df, "../output/mvt_wide_400_summary.rds")
mon <- monitor(fit)
mean(mon$Rhat>1.01)
mean(mon$Bulk_ESS<400)
mean(mon$Tail_ESS<400)
```

```{r}
N <- 10000
A <- diag(nrow=N)
model <- stan_model("mvt_wide.stan")
fit <- sampling(model, data=list(N=N), iter=1000, chains=4)
saveRDS(fit, "../output/mvt_wide_1000.rds")

full_data <- rstan::extract(fit, permuted=F, inc_warmup=F)
nparams <- dim(full_data)[3]
m_flattened <- matrix(nrow = nrow(full_data) * 4, ncol = (nparams + 1))
k <- 1
for(i in 1:4){
  for(j in 1:nrow(full_data)){
    m_flattened[k, 1:nparams] <- full_data[j, i, ]
    m_flattened[k, (nparams + 1)] <- i
    k <- k + 1
  }
}
m_flattened <- m_flattened %>% 
  as.data.frame()
colnames(m_flattened)[nparams + 1] <- "chain"
m_flattened <- m_flattened %>% 
  mutate(chain=as.factor(chain))
n_distinct(m_flattened$chain)

a_df <- a_df %>%
  tibble(samples_1000=colMeans(mAccuracy)/(1/n_distinct(training_data$chain)))
saveRDS(a_df, "../output/mvt_wide_summaries.rds")
mon <- monitor(fit)
mean(mon$Rhat>1.01)
mean(mon$Bulk_ESS<400)
mean(mon$Tail_ESS<400)

g <- 
  a_df %>% 
  melt() %>% 
  ggplot(aes(x=value, fill=as.factor(variable))) +
  geom_histogram(position="identity", alpha=0.8, bins=20) +
  scale_fill_grey("Model", labels=c("400", "1000")) +
  xlab(TeX("$R*$")) +
  ylab("Count") +
  geom_vline(xintercept = 1, linetype=2) +
  theme(text = element_text(size=14, colour="black"),
        legend.position = c(0.8, 0.8))
ggsave("../output/very_wide_data.pdf", g, width = 12, height = 8)
```
